Optimal design of experiment

### 2024-01-21 SFE_article_Design_of_Experiment/Article_Design_of_Experiment.tex
21.01

% ===================================================
% METADATA
% ===================================================
\title[mode=title]{Sensitivity Analysis}                      
\shorttitle{Sensitivity Analysis}

@@

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
This study aimed to investigate the supercritical extraction process of caraway oil from caraway seeds. The extraction was performed in a partially filled extractor with a fixed bed operated under multiple operating conditions. A distributed-parameter model describes the fluid-solid extraction process with $CO_2$ as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The local sensitivity analysis investigates the influence of infinitely small changes in the inlet temperature, pressure, and flow rate on the extraction yield.

@@

\begin{keywords}
Supercritical extraction \sep Sensitivity analysis \sep Mathematical modelling
\end{keywords}

@@

\subsection{Sensitivity Analysis} \label{CH: Sensitivity_Analysis}
\subfile{Sections/Sensitivity_Analysis}

@@

\section{Results}
\subfile{Sections/Results_Sensitivity}

@@

The sensitivity analysis is a tool to understand the parametric dependence of dynamic behavior and its suitability in inferring mechanisms of the dynamic behavior of the analyzed system. The mathematical formulation involves derivative-based local sensitivity analysis of the solution of an inverse problem with respect to perturbations of parameters. The local method is based only on considering all parameters as constants (those independent of time or location), and then the sensitivity equations are estimated by solving a variational system simultaneously with the original system. The sensitivity equations can be obtained in various ways, but the automatic differentiation technique was implemented to obtain the sensitivity equations in this work. By identifying which parameters are influential at each period of the simulation, local sensitivity analysis can yield valuable information for guiding the process modelling, the design of experiments or the model reduction. The local sensitivity analysis techniques consider the sensitivities at only a small region of parameter space, and the conclusions derived from such an analysis are limited to local conditions unless the discussed system is a linear system.

In this study, the local sensitivity analysis method was introduced to explore the dynamics of the supercritical extraction system, which consists of a set of partial differential equations. The sensitivity equations evaluate the influence of control variables such as flow rate, pressure, and inlet temperature on the state-space consisting of the concentration of the solute in solid and liquid phases, the fluid's enthalpy density, system pressure, and extraction yield. Every change in the control variables leads to a state-space change, eventually affecting the process performance. This information can be utilized to identify which control variables influence the extraction yield the most. The controls with high sensitivities on the extraction can be used to investigate to find optimal operating conditions from an economic point of view.

>>>>

% ===================================================
% METADATA
% ===================================================
\title[mode=title]{Optimal design of experiment}                      
\shorttitle{Optimal design of experiment}

@@

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
This study aimed to investigate the supercritical extraction process of caraway oil from caraway seeds. The extraction was performed in a partially filled extractor with a fixed bed operated under multiple operating conditions. A distributed-parameter model describes the fluid-solid extraction process with $CO_2$ as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. 
The optimal design of experiment problem is solved to find an experimental strategy for validating the correlations between extraction kinetic parameters and the operating conditions. This work proposes two experimental strategies: one that only manipulates the inlet temperature and the second that manipulate only the pressure.

@@

\begin{keywords}
Supercritical extraction \sep Optimal design of experiment \sep Mathematical modelling \sep Optimization \sep Fisher information
\end{keywords}

@@

\subsection{Design of experiment} \label{CH: Optimal_design_of_experiment}
\subfile{Sections/Optimal_design_of_experiment}

@@

\section{Results}
%\subfile{Sections/Results_Sensitivity}

@@

### 2024-01-21 SFE_article_Design_of_Experiment/Sections/IC_BC.tex
21.01

\documentclass[../Article_Model_Parameters.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

@@

	During the preparation period, when the operating conditions are achieved, the solute diffuses to the fluid phase in contact with the solid particles. Later, the solute in the fluid phase is partially moved (if the pressure increase in the system, the pump cause the movement of the fluid, even if the outlet valve is closed) to the region where there is no solid phase. As a result, the distribution of solutes mass in the fluid phase is assumed to not be uniform, and described by an arbitrary function $H=H(z)$. Some conclusions can be drawn from the analysis of the initial part of each yield curve. It can be noticed that each curve at the beginning has a curvature, which is not linear. In a general sense, it can be said that a quadratic function could approximate the initial part of each extraction curve. A function that, after integration, gives a quadratic-like result is a straight line. Based on that observation, the solute concentration in the fluid phase is assumed to be linearly distributed. The solute concentration is assumed to be zero at the outlet and reach the maximum at the beginning of the fixed bed. The graphical representation of the solute concentration in the fluid phase is shown in Figure \ref{fig: C_f_linear}.

>>>>

\documentclass[../Article_Design_of_Experiment.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

@@

	During the preparation period, the solute diffuses to the fluid phase in contact with the solid particles although the operating conditions are not obtained yet. Later, the solute in the fluid phase is partially moved (if the pressure increase in the system, the pump cause the movement of the fluid, even if the outlet valve is closed) to the region where there is no solid phase. As a result, the distribution of solutes mass in the fluid phase is assumed to not be uniform, and described by an arbitrary function $H=H(z)$. Some conclusions can be drawn from the analysis of the initial part of each yield curve. It can be noticed that each curve at the beginning has a curvature, which is not linear. In a general sense, it can be said that a quadratic function could approximate the initial part of each extraction curve. A function that, after integration, gives a quadratic-like result is a straight line. Based on that observation, the solute concentration in the fluid phase is assumed to be linearly distributed. The solute concentration is assumed to be zero at the outlet and reach the maximum at the beginning of the fixed bed. The graphical representation of the solute concentration in the fluid phase is shown in Figure \ref{fig: C_f_linear}.

### 2024-01-21 SFE_article_Design_of_Experiment/Sections/Literature_Review.tex
21.01

\documentclass[../Article_Model_Parameters.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	This study aims to analyze the influence of changes in operating conditions on the supercritical extraction process described in {\color{red}article 1}. The emphasis is put on the effect of the mass flow rate, pressure, and the inlet temperature. The relation between input and output is obtained by applying a sensitivity analysis. Sensitivity analysis examines the impact of varying inputs or model parameters on the uncertainty of a mathematical model or system's output. The aim is to understand and to allocat the source of uncertainty in the output to the corresponding inputs or parameters. There are many sensitivity analysis methods, which include but are not limited to those listed below:

@@

	Different supercritical extraction models were analyzed using sensitivity analysis. One example of such work is \citet{Fiori_2007}. The sensitivity calculations were performed by varying the parameters within their confidence interval and observing how the model results changed. This allows to establish the effect of the uncertainties on model predictions. The sensitivity analysis revealed that the particle diameter and the internal mass transfer coefficient are significant for the extraction process. The effect of changing some operative conditions was also investigated, underlining how the solvent flow rate and the seed milling affect the extraction process.

>>>>

\documentclass[../Article_Design_of_Experiment.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	This study aims to analyze the influence of changes in operating conditions on the supercritical extraction process described in {\color{red}article 1}. The emphasis is put on the effect of the mass flow rate, pressure, and the inlet temperature. The relation between input and output is obtained by applying a sensitivity analysis. Sensitivity analysis examines the impact of varying inputs or model parameters on the system's output. The aim is to understand and to allocate the source of uncertainty in the output to the corresponding inputs or parameters. There are many sensitivity analysis methods, which include but are not limited to those listed below:

@@

	Different supercritical extraction models were analyzed using sensitivity analysis. \citet{Fiori_2007}, performed the sensitivity calculations by varying the parameters within their confidence interval and observing how the model results changed. This allows to evaluate the effect of the uncertainties on model predictions. The sensitivity analysis revealed that the particle diameter and the internal mass transfer coefficient are significant for the extraction process. The effect of changing some operative conditions was also investigated, underlining how the solvent flow rate and the seed milling affect the extraction process.

### 2024-01-21 SFE_article_Design_of_Experiment/Sections/Low_mach_number_expansion.tex
21.01

\documentclass[../Article_Model_Parameters.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	As discussed by \citet{Lions2013}, the low Mach number equations are a subset of the fully compressible equations of motion (continuity, momentum and energy). Such a set of equations allow for large variations in gas density but it is consider to be acoustically incompressible. The low Mach number equations are preferred over the full compressible equations for low speed flow problems $\left( M_a = \frac{|V|}{\sqrt{ \partial P / \partial \rho }} \ll 1 \right)$ to avoids the need to resolve fast-moving acoustic signals. The equations are derived from the compressible equations based on the perturbation theory. The perturbation theory develops an expression for the desired solution in terms of a formal power series known as a perturbation series in some "small" parameter $\zeta$, that quantifies the deviation from the exactly solvable problem. The leading term in this power series is the solution of the exactly solvable problem, while further terms describe the deviation in the solution, due to the deviation from the initial problem.

>>>>

\documentclass[../Article_Design_of_Experiment.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	As discussed by \citet{Lions2013}, the low Mach number equations are a subset of the fully compressible equations of motion (continuity, momentum and energy). Such a set of equations allow for large variations in gas density but it is consider to be acoustically incompressible. The low Mach number equations are preferred over the full compressible equations for low speed flow problems $\left( M_a = \frac{|v|}{\sqrt{ \partial P / \partial \rho_f }} \ll 1 \right)$ to avoids the need to resolve fast-moving acoustic signals. The equations are derived from the compressible equations based on the perturbation theory. The perturbation theory develops an expression for the desired solution in terms of a formal power series known as a perturbation series in some "small" parameter $\zeta$, that quantifies the deviation from the exactly solvable problem. The leading term in this power series is the solution of the exactly solvable problem, while further terms describe the deviation in the solution, due to the deviation from the initial problem.

### 2024-01-21 SFE_article_Design_of_Experiment/Sections/Low_mach_number_expansion_imp.tex
21.01

	The equations \ref{EQ: CompressibleEuler_1} to \ref{EQ: CompressibleEuler_2} describe the fully compressible equations of motion (respectively, the transport of mass, momentum and energy) for the quasi-one-dimensional case. We rescale the time variable, considering finally

@@

	In one-dimensional case, the equation \ref{EQ: Incompressibility_Condition_Ideal_IsoThermal} become equivalent of $\text{div} ( A_f v) = 0$, which known as the incompressibility condition (\citet{Lions2013}).

>>>>

	The Equations \ref{EQ: CompressibleEuler_1} to \ref{EQ: CompressibleEuler_2} describe the fully compressible equations of motion (respectively, the transport of mass, momentum and energy) for the quasi-one-dimensional case. We rescale the time variable, considering finally

@@

	In one-dimensional case, the Equation \ref{EQ: Incompressibility_Condition_Ideal_IsoThermal} become equivalent of $\text{div} ( A_f v) = 0$, which known as the incompressibility condition (\citet{Lions2013}).
